\thetitle

### 2020-11-14 writing/rational.tex
Move debugging scenarios description into previous section
format + prose tweaks
revising related work

Based on this idea, a metric for migratory paths is a concise way to
evaluate the effectiveness of a rational programmer. The components of
an ill-typed program, such as one of  our ill-typed mutants, 
create a \emph{debugging scenario space}, a lattice like
the one introduced by ~\citet{tfgnvf-popl-2016} for performance
evaluations. As we discuss briefly at the end of ~\ref{sec:mutate}, 
this space consists of variants of the program, dubbed
\emph{scenarios}, that differ in their type specifications. In the
setting of migratory typing,\footnote{The method can be adapted to
so-called true gradual typing systems~\cite{svcb-snapl-2015} in a
straightforward manner.} each scenario differs from another in which
components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is a subset of
$\set{\component}$ . These scenarios of $\system$ are ordered by the
subset relation and thus form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements.  The bottom scenario of
$\lattice{\system}$ is always $\emptyset$ and the top one is $\system$
itself. In between are all the mix-typed scenarios.

@@

When the buggy untyped component of a program is replaced by the typed counterpart,
the type checker fails if this component causes the impedance mismatch. Hence, a
trail that end at an ill-typed scenarios successfully pinpoints the location of
bugs. 
\begin{quote}
\it A natural blame trail $\conf_0,...\conf_n$ in a lattice $\lattice{\system}$ is
\emph{successful} iff its last scenario $\conf_n$ does not type check.  A natural
blame trail $\conf_0,..,\conf_n$ in a lattice $\lattice{\system}$ is \emph{failing}
iff $\conf_n$ type checks and the trail cannot be extended further.
\end{quote}
That is failing natural blame trails are those that have not reached a scenario that
reveals the bug at compile time and their last scenario does not produce blame. Thus
the rational programmer receives no hints from the gradual type system on how to
continue the search for the bug.

@@

Even if the success of the natural rational programmer is not by pure chance, it
could well be because the natural semantics perform runtime type checks at the same

@@

In detail we answer $Q_1$ for an ill-typed mutant from
section~\ref{sec:mutate} by creating trails for each natural blame,
natural exception and natural random trail in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and determine the scenario that can detect the fault in
the program. This can happen either because Typed Racket's type checker outright
rejects the scenario or because natural Typed Racket's runtime performs a type check
that fails. In either case, each of these scenarios is the root of a trail. Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail---a trail may the top of the finite lattice---we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
roots two more times; once following the natural-exception programmer and once the
natural-random programmer. Finally, based on the collected data, we calculate the
success/failure results for the trails to determine whether natural blame is more
successful than natural exceptions and pure chance and we obtain the answer for
$Q_1$.  The process is analogous for the other questions employing the respective
modes of the rational programmer. Figure~\ref{fig:process} summarizes the
experimental process for one mode of the rational programmer and connects
it with the mutations from section~\ref{sec:mutation}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{./Images/process}
  \caption{The experimental process for one mode of the rational
  programmer}
  \label{fig:process}
\end{figure}

>>>>

Based on this idea, a metric for migratory paths is a concise way to
evaluate the effectiveness of a rational programmer. The components of
an ill-typed program, such as one of  our ill-typed mutants, 
create a \emph{debugging scenario space}, a lattice like
the one introduced by ~\citet{tfgnvf-popl-2016} for performance
evaluations. As we discuss briefly at the end of section ~\ref{sec:mutate}, 
this space consists of variants of the program, dubbed
\emph{scenarios}, that differ in their type annotations. In the
setting of migratory typing,\footnote{The method can be adapted to
so-called true gradual typing systems~\cite{svcb-snapl-2015} in a
straightforward manner.} each scenario differs from another in which
components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is a subset of
$\set{\component}$ . These scenarios of $\system$ are ordered by the
subset relation and thus form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements.  The bottom scenario of
$\lattice{\system}$ is always $\emptyset$ and the top one is $\system$
itself. In between are all the mix-typed scenarios.

@@

When the buggy untyped component of a program is replaced by the typed counterpart,
the type checker fails if this component causes the impedance mismatch. Hence, a
trail that ensd at an ill-typed scenarios successfully pinpoints the location of
bugs. 
\begin{quote}
\it A natural blame trail $\conf_0,...\conf_n$ in a lattice $\lattice{\system}$ is
\emph{successful} iff its last scenario $\conf_n$ does not type check.  A natural
blame trail $\conf_0,..,\conf_n$ in a lattice $\lattice{\system}$ is \emph{failing}
iff $\conf_n$ type checks and the trail cannot be extended further.
\end{quote}
That is failing natural blame trails are those that have not reached a scenario that
reveals the bug at compile time and their last scenario does not produce blame. Thus
the rational programmer receives no hints from the gradual type system on how to
continue the search for the bug.

@@

@@

In detail we answer $Q_1$ for an ill-typed mutant from
section~\ref{sec:mutate} by creating trails for each natural blame,
natural exception and natural random trail in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and select an interesting debugging scenario
as we discuss in section~\red{sec:mutate} to be the root of a trail. Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail---a trail may the top of the finite lattice---we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
roots two more times; once following the natural-exception programmer and once the
natural-random programmer. Finally, based on the collected data, we calculate the
success/failure results for the trails to determine whether natural blame is more
successful than natural exceptions and pure chance and we obtain the answer for
$Q_1$.  The process is analogous for the other questions employing the respective
modes of the rational programmer. Figure~\ref{fig:process} summarizes the
experimental process for one mode of the rational programmer and connects
it with the mutations from section~\ref{sec:mutation}.

\begin{figure}
  \centering
  \includegraphics[scale=0.36]{./Images/process}
  \caption{The experimental process for one mode of the rational
  programmer}
  \label{fig:process}
\end{figure}

### 2020-11-14 writing/related.tex
revising related work

Besides the three semantics we consider, the literature on gradual typing
comes with a plethora of different approaches. For example, Pyret
(\shorturl{https://www.}{pyret.org}) treats function types same as Natural
and data types same as Erasure. The Forgetful~\cite{cl-icfp-2017} and the
Amnesic~\cite{gfd-oopsla-2019} semantics are similar  to Transient but
use wrappers instead of inlined checks.  Nom~\cite{mt-oopsla-2017} and
other \emph{concrete\/} semantics~\cite{wnlov-popl-2010, rsfbv-popl-2015,
mt-oopsla-2017, rzv-ecoop-2015, rat-oopsla-2017} assume that every type
maps to a unique known tag and they perform type tag checks  for every
value that crosses from typed to untyped parts of a program or vice versa.
With the exception of Amnesic~\cite{gfd-oopsla-2019}, all the above
semantics forego blame. However, Amnesic is a mere
theoretical construction and is not meant as a blueprint for an
implementation.

>>>>

Besides the three semantics we consider, the literature on gradual typing
comes with a plethora of different approaches. For example, Pyret
(\shorturl{https://www.}{pyret.org}) treats function types same as Natural
and data types same as Erasure. The Forgetful~\cite{cl-icfp-2017} and
the Amnesic~\cite{gfd-oopsla-2019} semantics are similar  to Transient but
use wrappers instead of inlined checks.  Nom~\cite{mt-oopsla-2017} and
other \emph{concrete\/} semantics~\cite{wnlov-popl-2010, rsfbv-popl-2015,
rzv-ecoop-2015, rat-oopsla-2017} assume that every type maps to a unique
known tag and they perform type tag checks  for every value that crosses
from typed to untyped parts of a program or vice versa.  Monotonic
references~\cite{svctg-esop-2015} and the semantics~\cite{tlt-popl-2019,
etg-icfp-19, tt-scp-20, tgt-popl-18, tt-sas-17} derived with the
"turn-the-crank-get-a-POPL-paper" Abstracting Gradual Typing
technique~\cite{gct-popl-2016} are (optimized) variants of Natural for
different language features.  With the exception of Amnesic and Nom, all
the above semantics forego blame. However,  Nom's gradual types impose
significant restrictions on programmers, while Amnesic is a mere
theoretical construction and is not meant as a blueprint for an
implementation.
